---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pseudonymizer

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/pseudonymizer)](https://CRAN.R-project.org/package=pseudonymizer)
<!-- badges: end -->

The goal of `pseudonymizer` is to pseudonymize sensitive data. The package can be used in cases where a research team receives or collects sensitive data that has to be pseudonymized before it can be analyzed. The package was built for a specific use case, but it can be used as a template to pseudonymize other data sets.

The idea is to designate a data security manager, who is not part of the research team and who has the task to securely store the original data and to pseudonymize the data. `pseudonymizer` is the tool through which the research team and the data security manager can collaborate. For example, if the researchers need month and year of birth for their analysis, they can adapt the package to extract this information from the date of birth. The data security manager can then apply the package and provide the research team the pseudonymized data.

The package involves six steps:

1) Import a data set
2) Generate a unique, stable, pseudonymized identifier from a sensitive identifier for each observation unit  
3) Aggregate some sensitive data (like aggregate date of birth to year of birth)
4) Optional: Generate a keytable which holds the sensitive and the pseudonymized identifier
5) Optional: Generate an address file
6) Drop sensitive data and export


Some points are noteworthy:

1) Which (combination of) variables are considered sensitive depends on different factors and should be defined in accordance with any existing data protection contract.
2) This package generates a pseudonymized ID using SHA-256, a vectorized hash function which is widely used, particularly by [U.S. government agencies to secure their sensitive data](https://csrc.nist.gov/publications/detail/fips/180/4/final). A hash is a one-way mathematical function (i.e. it can't be reverse engineered). However, it is in principle possible to generate large data bases of possible inputs and compute the corresponding hash (especially if the structure of the input is known). With such a key table, re-identification would be possible. Adding a constant value (the 'salt') to all inputs increases security substantially. 
3) Choose a salt that is long and completely unrelated to the data or the project, ideally generated randomly.
4) The key table with the identifying data and the salt must only be accessible by the data security manager and must be stored as securely as the original data.
5) Through the combination of several variables, pseudonymized data might still allow the identification of certain individuals. Therefore, this data should also be stored securely and should only be accessible to a restricted and authorized group of persons. 

## Installation

You can install the development version of pseudonymizer from [GitHub](https://github.com/) with:

``` r
devtools::install_github("KristinaSchuep/pseudonymizer")
```

## Example

This is a basic example with some fake data. The function `pseudonymize` performs all of the six steps described above and exports a keytable, an address file and the final, pseudonymized data set ready for distribution to the research team.

```{r example, eval = TRUE}
library(pseudonymizer)

# Load your salt
# Beware, this is NOT a good salt, this is for illustrative purposes only!
mysalt <- "myverygoodsalt" 
# Use this code to import your own, safe salt:
# mysalt <- as.character(read.delim("salt.txt",header=FALSE, sep = "", dec="."))

# Define variables to rename
oldnames <- c("NNSS","NACHNAME","VORNAME","GEBURTSDATUM")
newnames <-  c("ahvnr","surname","firstname","birthday")

# Define variables to export for address table
address_vars <- c('pseudo_id', 'firstname', 'surname', 'strasse', 'hausnr','plz', 'wohnort')

# Define variables to export for keytable
keytable_vars <- c("ahvnr","firstname","surname","birthday","strasse","hausnr", "plz", "wohnort")

# Define sensitive variables that should not be exported for the final data set
sensitive_vars <- c("ahvnr","firstname","surname","birthday","strasse","hausnr", "plz", "wohnort")

# Perform all steps in one
df <- pseudonymize(data_name = "FAKE_DATA_2019",
                   import_filename = "data-raw/FAKE_DATA_2019.xlsx", 
                   import_oldnames = oldnames,
                   import_newnames = newnames,
                   id_original = "ahvnr",
                   id_salt = mysalt,
                   export_path = "output",
                   address_vars =  address_vars,
                   keytable_vars = keytable_vars,
                   sensitive_vars = sensitive_vars)
```

Instead of using this wrapper function, you can also perform each step separately:

```{r}
# Step 1: Import dataset into R
df <- import_raw_data(filename = "./data-raw/FAKE_DATA_2019.xlsx",
                      oldnames = oldnames,
                      newnames = newnames)

# Step 2: Generate unique ID from AHV-number 
df <- unique_id(data = df, id = "ahvnr", salt = mysalt)

# Step 3: Aggregate sensitive data 
df <- aggregate_sensitive(data = df)

# Step 4: Generate and append key table 
append_keytable(df = df,
                path = "output",
                keytable_vars = keytable_vars,
                id_original = "ahvnr")

# Step 5: Generate address file 
export_address(data = df,
               path = "output",
               data_name = 'FAKE_DATA',
               vars = address_vars)

# Step 6: Drop sensitive data and export 
df <- export_pseudonymized(data = df,
                           sensitive_vars = sensitive_vars,
                           path = "output",
                           data_name = "FAKE_DATA",
                           data_summary = TRUE)
```


